---
title: "Noodling on LD decay"
author: "Maya Wilson Brown"
date: "2024-10-16"
output: html_document
---


First, the script that generated the calculations of LD
based on population distinctions
```{r}
pops <- read.csv("~/Documents/PhD/Research/capsella_population_structure/cbp_pop_str.txt", sep = "\t", header=T)

nyc <- pops[which(str_detect(pops$vcf_sample_name, "L003")),]
easia <- pops[which(pops$k3population == "E_Asia"),]

write.table(nyc$vcf_sample_name, "~/Documents/PhD/Research/vcf_filtering/ld_decay/nyc_vcf_sample.txt",
            row.names = F,
            col.names = F,
            quote = F)

write.table(easia$vcf_sample_name, "~/Documents/PhD/Research/vcf_filtering/ld_decay/easia_vcf_sample.txt",
            row.names = F,
            col.names = F,
            quote = F)
```


```{bash, code = readLines("~/Documents/PhD/Research/vcf_filtering/ld_decay/noodling_LD_decay.sh"), eval = FALSE}
```

The actual calculation of linkage into averaged bins is done by the following python script.
```{bash, code = readLines("~/Documents/PhD/Research/vcf_filtering/ld_decay/ld_decay_calc.py"), eval = FALSE}
```

Then, to plot the output
```{r}
library(ggplot2)
# read in data
ld_bins <- read.csv("~/Documents/PhD/Research/vcf_filtering/ld_decay/nyc_chrAll.ld_decay_bins",
                    sep = "\t",
                    header = T)


# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2), colour = chr) + geom_line() +
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  geom_hline(yintercept = 0.1, color = "red") +
  facet_wrap(~chr) +
  theme_bw()

# which bins size before LD reaches backgound level of about 0.1
ggplot(subset(ld_bins, avg_R2 <= 0.1), aes(distance, avg_R2), colour = chr) + geom_line() +
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  geom_vline(xintercept = 500000, color = "red") +
  facet_wrap(~chr) +
  theme_bw()

ggplot(ld_bins, aes(distance, avg_R2), colour = chr) + geom_line() +
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  geom_vline(xintercept = 500000, color = "red") +
  geom_hline(yintercept = 0.1, color = "blue") +
  facet_wrap(~chr) +
  theme_bw()
```

Honestly, 500kb seems kind of long. I wonder if introgression from C. rubella might have something to do with it, so I will do the same analysis on the East Asian population.
```{r}

```

